Applying Deep Learning to Acoustic Analysis in the PICO-60 Dark Matter Detector

### 2018-08-09 documentation/paper.tex
Clarified grav diff quite a bit

Gravitational differentiation is a novel technique for gradient computation in the final layer of a neural network. It is intended for training a neural network in a semi-supervised fashion on a relatively small set of imperfectly labeled data, while simultaneously incorporating the network's changing predictions on a set of unlabeled data to encourage decisive classifications.

@@

It makes use of the gravitational differentiation function $GravDiff(p, \psi, g)$, which is parameterized by the network's existing prediction $p$ on the training example in question, the degree $\psi$ of the piecewise exponential function used to distort the response of the gradient, and the gravitational multiplier $g$. The function is defined as follows:

@@

The essential purpose of $GravDiff(p, \psi, g)$ is to allow confidently classified (near 0 or 1) unlabeled examples to influence the training process, pushing them further towards the classification they are already near (like gravity), while ensuring examples with low-confidence (near 0.5) classifications do not have a significant effect.

@@

However, what is really desired is a shallow slope in the center (so low-confidence network outputs in the range of, for instance, 0.2 to 0.8 will produce very shallow output gradients). Applying a relatively large exponent (on the order of 9) accomplishes this, because values less than 1 will rapidly approach 0. However, this only works with odd integer exponents (because they produce negative outputs for negative inputs), limiting the ability to fine-tune this function by changing $\psi$. To resolve this, a piecewise exponential function is used, where the absolute value is taken prior to applying the power and the sign is multiplied in afterwards. This permits use of the full range of curves produced by even and non-integer values of $\psi$.

>>>>

Gravitational differentiation is a novel technique for training a neural network in a semi-supervised fashion on a relatively small set of imperfectly labeled data, while simultaneously incorporating the network's changing predictions on a set of unlabeled data to encourage decisive classifications.

It accomplishes this using a new method for calculating gradients for the last layer of the neural network. Based on the current predictions of a trained network on a certain unlabeled training example, a gradient is calculated for the last layer with the following purposes:

\begin{itemize}
    \item For examples with confident predictions close to 0 or 1: Create a large gradient that pulls the prediction closer to 0 or 1 through gradient descent (like gravity).
    \item For examples with low-confidence predictions close to 0.5: Create a gradient near 0 that does not affect training significantly.
\end{itemize}

The gradients produced by this technique are used for training in the same way as gradients calculated with a loss function. Backpropagation is used to calculate gradients for previous layers, and a stochastic gradient descent optimizer is used for weight updates.

@@

Gradients are calculated using the gravitational differentiation function $GravDiff(p, \psi, g)$, which is parameterized by the network's existing prediction $p$ on the training example in question, the degree $\psi$ of the piecewise exponential function used to distort the response of the gradient, and the gravitational multiplier $g$. The function is defined as follows:

@@

@@

However, what is really desired is a shallow slope in the center (so low-confidence network outputs near 0.5 will produce very shallow output gradients). Applying a relatively large exponent (in the range of 3 to 11) accomplishes this; it squashes values close to 0 closer to 0.

However, this only works with odd integer exponents (because they produce negative outputs for negative inputs), limiting the ability to fine-tune this function by changing $\psi$. To resolve this, a piecewise exponential function is used, where the absolute value is taken prior to applying the power and the sign is multiplied back in afterwards. This permits use of the full range of curves produced by even and non-integer values of $\psi$.
